---
title: "Introduction to gRumble"
author: "MBayOtolith <jmoxley@mbayaq.org>"
date: "May 2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gRumble}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The gRumble package contains functions for the post-processing of sensor data returned from animal-borne motion detecting data loggers (a.k.a., inertial monitoring units or IMUs, for short) with shifting orientations during deployment. 


Inertial measurement data (e.g., accelerometry, magnetometry, etc.) from such data loggers are transforming quantitative understanding of animal behavior, but requires significant and careful handling to translate high-resolution data streams into interpretible behavior.  One primary post-processing task is repositioning the tri-axial sensor data so that measurements correspond to the body posture of the animal.  Aligningment of the data to the animal's frame of reference is accomplished by resolving the static acceleration signal of gravity and/or aligning dynamic acceleration signals to known (and ideally repetetive) behaviors.  

![This figure depicts the differences in data when using an unfixed sensor ("stomach") and a fixed sensor ("fin mount")](paste(getwd(), "scratch/stomch_vs_fin.png", sep = "/"))


Here we present a dynamic approach to transforming raw IMU data in order to align signals with the body axes of a tagged animal, in this case sharks carrying unfixed stomach-borne IMUs fed to them during research.  This approach builds on methods described in [Jorgensen et al. 2015] (https://animalbiotelemetry.biomedcentral.com/articles/10.1186/s40317-015-0071-6), but importantly uses a dyanmic moving window instead of a discrete window that allows for continuous correction and accounting for stomach movement.  

## History of gRumble package

Connor is a badass, and just straight coded up this fine piece of programming in 2017.  But it builds off the slick methods outlined in Jorgensen et al. 2015.  

## Model overview
The correction procedure follows four primary steps, three of which are accomplished by custom functions: 

1.  findBreaks() - indexing function to identifying qualifying periods of horizontal/prone behavior where the static acceleration of gravity can be isolated;
2.  EstAngles() - estimation function to calculate the X/Y/Z angles within each window of horizontal/prone behavior that can correct data to the animal's frame of reference;  
3.  Matching of the anterior/posterior axes - ensuring alignment of the anterior/posterior axes of the animal and the IMU;
4.  angInterp() - interpolation function to backfill estimated corrections calculated at each window's midpoint to the entire data range stream.  

Notes about data preparation and smoothing pre-processing are discussed in a separate section below.  In our example, the procedure relies on assumptions that the shark's body posture is horizontal or prone during periods when the vertical velocity (or depth rate) is near zero.  When adapting to other species, care should be taken in tailoring this assumption to appropriate behaviors for that animal.  

```{r, echo = F, message = F}
#load necessary resources
library(devtools)

#pull down latest version & load
install_github("MBayOtolith/gRumble/gRumble"); library(gRumble);

#adjust working dir
wd <- "/Users/jhmoxley/Documents/Biologia & Animales/[[SharkTank]]/GitTank/gRumble"
setwd(wd)

#load data
dat <- read.csv("6_Pepe_20101105_clipped.csv", header = T)
head(dat); str(dat)
datFreq <- 5;    #sampling frequency of data

####DATA PROCESSING
dat <- dat[1:(4*3600*datFreq),]    #working data, subset for first 4 hrs

#smooth depth (in this case over 5s)
dat$Depth <- filter(dat$Depth, filter = rep(1,5*datFreq)/(5*datFreq), sides = 2, circular = T)

#calc vertical velocity
dat$VV <- c(0, diff(dat$Depth, 1))
dat$VV <- filter(dat$VV, filter=rep(1,datFreq), sides=2, circular=TRUE)  #converts/smooths to m/s from m/.2s
```

## findBreaks()
The gRumble method depends on the ability to identify moments of prone/horizontal posture when a properly-oriented tag will record static acceleration along a single axis.  For a swimming animal like a shark, this is determined by the animal's vertical velocity or depth rate as it is recorded by the depth sensor.  Applications to other species can define similar behaviors of known body posture from appropriate sensors to accomplish similar corrections.  

```{r, echo = T}
#finds windows in VV-vector over which animal has "horizontal/prone" posture
brks <- findBreaks(VV = dat$VV, corrFreq = 30, minsize = 10, lims = c(-0.5, 0.5), dat_Freq = 5)
```

This function call inputs the vector of vertical velocities from the IMU depth/pressure sensor and a range of arguements that structure the correction/estimation procedure.  The arguement *corrFreq* how often (in seconds) to estimate a correction, while *minsize* sets the lower limit for an adequate window to average VV across and test if the mean lies within the thresholds set by *lims*.  The arguement *dat_Freq* is set by the temporal resolution of the input data, and works with *corrFreq* to structure how the windows expand to find periods when VV averages towards 0.  

**FIGURE ON WINDOW DEFINITION IN A DATA STREAM**
**see figure on proper smoothing window?? **

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## Angle calculation
Description of EstAngs()
2 steps.. matching methods for estimating X/Y; 
mismathcing method for finding Z (requires understanding of signal param)

##Correction for anterior/posterior 

##Backfill angInterp()

##apply corrections rotAngle()

##notes on data/animal, data formats, and pre-processing
-designed around procedures needed in shark research, applicable other creatures
-test data is XX hours from a research shark
-dependencies
-smoothing & filtering, and defining the depth rate

##tidbits about code
You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
