---
title: "Introduction to gRumble"
author: "Jerry Moxley"
date: "May 2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gRumble}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The gRumble package contains functions for the post-processing of sensor data returned from motion detecting data loggers (a.k.a., inertial monitoring units or IMUs, for short) with shifting orientations during deployment.   

Accelerometry data from such data loggers are transforming quantitative understanding of animal behavior, but requires significant and careful handling to translate high-resolution data streams into interpretible behavior.  A primary task of post-processing this data is to calibrate the tag's orientation so that tri-axial measurements correspond with the body posture of the animal.  This is primarily accomplished by resolving static acceleration due to the force of gravity and re-orienting the data with coordinate transformations to minimize 

This package was designed for stomach-borne telemetry devices fed to sharks, but applies to other IMU uses that involve unknown orientations during the deployment.  


## History of gRumble package

Connor is a badass, and just straight coded up this fine piece of programming in 2017.  But it builds off the slick methods outlined in Jorgensen et al. 2015

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
