---
title: "Introduction to gRumble"
author: "MBayOtolith <jmoxley@mbayaq.org>"
date: "May 2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gRumble}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The gRumble package contains functions for the post-processing of sensor data returned from animal-borne motion detecting data loggers (a.k.a., inertial monitoring units or IMUs, for short) with shifting orientations during deployment. 


Inertial measurement data (e.g., accelerometry, magnetometry, etc.) from such data loggers are transforming quantitative understanding of animal behavior, but requires significant and careful handling to translate high-resolution data streams into interpretible behavior.  One primary post-processing task is repositioning the tri-axial sensor data so that measurements correspond to the body posture of the animal.  Aligningment of the data to the animal's frame of reference is accomplished by resolving the static acceleration signal of gravity and/or aligning dynamic acceleration signals to known (and ideally repetetive) behaviors.  


Here we present a dynamic approach to transforming raw IMU data in order to align signals with the body axes of a tagged animal, in this case sharks carrying unfixed stomach-borne IMUs fed to them during research.  This approach builds on methods described in [Jorgensen et al. 2015] (https://animalbiotelemetry.biomedcentral.com/articles/10.1186/s40317-015-0071-6), but importantly uses a dyanmic moving window instead of a discrete window that allows for continuous correction and accounting for stomach movement.  

## History of gRumble package

Connor is a badass, and just straight coded up this fine piece of programming in 2017.  But it builds off the slick methods outlined in Jorgensen et al. 2015.  

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
